---
title: "Assignment 4 CRSS 8010"
author: "Matheus Ardigueri"
date: "09/15/2026"
output:
  html_document:
    df_print: paged
---

# 4 Assignment - ANOVA.

## A group is testing 3 different treatments for thrombocytopenia (low platelet count)

## The file 08_platelets.csv contains data from their trial of 96 patients (24 per treatment + controls)

### Opening File and gathering information.

```{r}
platelets <- read.csv("../data/08_platelets.csv")

platelets
```

### Extracting values for making easier creating the lists.

```{r}
unique(platelets$patient)

```

```{r}
unique(platelets$variety)
```

```{r}
platelets$platelet_count[platelets$variety == "Vitality"] # Change variety name to have values for that specific variety.
```

## Calculating manually the Analysis of Variance

```{r}
Vitality <- c(88.1,  74.4,  26.7,  22.0, -32.3,  63.5,  52.0,  80.3,  70.3,  75.3,  50.1,  63.1,  38.9, 128.0, 143.8, 126.1,  45.4,  94.4, 58.6,  88.6,  88.5, -49.7,  -0.4, -37.2)
Booster <- c(81.6 , 11.6 , 31.2 , 95.5 , 56.1,  93.4,  60.1, 124.8,  49.0, 166.3, -16.7,  52.1, 107.8, 103.8,  17.6, 106.0, -13.8,  16.7, 43.5 ,-53.7,  55.5,  96.1 , 24.5  ,45.7)
Protect <- c(153.3,  87.3,  91.8,  65.8,  84.5,  89.8,  92.1,  68.9,  92.4, 145.2, 108.3,  78.6,  98.8,  59.5,  45.3, 173.7, 151.6,  73.0,109.3, 102.7,  11.1,  70.1 ,114.7, 156.2)
none <- c(101.5, 147.3,  39.3,  79.1,  75.0, 192.5, 154.7, 176.6,  83.4, 225.6,  44.7, 121.7,   7.2,  24.7, 124.3,  25.1,  54.6,  46.6,17.8,  85.8, 133.5, 110.7, 170.6, 132.8)

scores <- c(Vitality, Booster, Protect, none)

#FIrst step is to calcuulate the mean

meanV <- mean(Vitality)
meanB <- mean(Booster)
meanP <- mean(Protect)
meanN <- mean(none)
mean_all <- mean(scores)

mean_all

# Second step is to calculate the sum of sqares

ssV <- (meanV - mean_all) ^2 * length(Vitality)
ssB <- (meanB - mean_all) ^2 * length(Booster)
ssP <- (meanP - mean_all) ^2 * length(Protect)
ssN <- (meanN - mean_all) ^2 * length(none)
ssall <- ssV + ssB + ssP + ssN

ssall
# Third step is to calculate the errors

errorV <- sum((Vitality - meanV)^2)
errorB <- sum((Booster- meanB)^2)
errorP <- sum((Protect- meanP)^2)
errorN <- sum((none- meanN)^2)
error <- errorV + errorB + errorP + errorN

error

dof_treatments <- 3 
dof_error <- 92

ms_trt <- ssall / dof_treatments 
ms_error <- error / dof_error 

F <- ms_trt / ms_error

F
```

## P-value

```{r}
Fstats2 <- 5.405009

# P-value

pf(Fstats2, df1 = 3, df2 = 92, lower.tail = FALSE)

```


## Calculating using AOV

```{r}
model <- aov(platelet_count ~ variety , data = platelets) # platelets is explained by the column treatment and the data is platelets

summary(model)
```

Results are the same for both calculations, meaning that I did the calculations right.
