---
title: "Untitled"
author: "Matheus Ardigueri"
date: "2026-10-06"
output: html_document
---

# Remembering exercises

```{r}
muscadines <- read.csv("./data/14_muscadines.csv")
muscadines

```

```{r}
library(ggplot2)

quality_control <- ggplot(data = muscadines, aes(
  x = block,
  y = yield,
  color = variety
)) +
  geom_jitter() +
  theme_bw()

quality_control

```

```{r}
quality_control2 <- ggplot(data = muscadines, aes(
  x = yield,
)) +
  geom_histogram() +
  facet_grid(~variety) +
  theme_bw()

quality_control2
```
```{r}
model <- aov(data = muscadines, yield ~ variety + block)

summary(model)
```
```{r}
hist(residuals(model), breaks = 20)

```
```{r}
plot(residuals(model), pch = 20, cex = 1.5)
```

```{r}
library(dplyr)

cleaned = muscadines %>%
  mutate(residuals = residuals(model))

cleaned

cleaned_keep <- cleaned %>%
  filter(residuals <= 5 | yield <= 12)

cleaned_keep
```

```{r}
hist(cleaned_keep$yield)

```
```{r}
plot(cleaned_keep$yield, pch = 20, cex = 1.5)

```

```{r}
model_clean <- aov(data = cleaned_keep, yield ~ variety + block)

summary(model_clean)
```
```{r}
hist(residuals(model_clean))
plot(residuals(model_clean), pch = 20, cex = 1.5)

```

```{r}
organic <- read.csv("./data/14_organic.csv")
organic
```

```{r}
model_organic <- aov(data = organic, organicmatter ~ biochar * compost)
summary(model_organic)
```
```{r}
interaction.plot(x.factor = organic$biochar,
                 trace.factor = organic$compost,
                  response = organic$organicmatter
                  )
```

# Linear regression

Shortcomes of anova
- Complex designs and missing data can make anova difficult.
- Didide data into groups and see which groups explain the variation
- Categorical variables
- Simple designs
- Requires normality



Linear regression - estimate the effect of different variables and check if they are different than 0, basically explain how well one or more variables can explain the data.
- Continuoues and categorical variables
- Pretty much any design
- Extended to any distribution of residuals

GOAL: Find the line that best summarizes a lot of data with minimal error.


```{r}
bn <- data.frame(
  plot = c(1,2,3,4),
biomass = c(17.6, 14.3 , 18.9, 12.6),
nitrogen = c(10,5,10,5)
  
  )

mymodel <- lm(biomass ~ nitrogen, data = bn)
summary(mymodel)
```

```{r}
mymodel2 <- aov(biomass ~ nitrogen, data = bn)
summary(mymodel2)
```
```{r}
yield <- read.csv("../data/14_yield.csv")
yield
```

```{r}
model_yield <- lm(yield ~ n + p + block, data = yield)
summary(model_yield)
anova(model_yield)
```

```{r}

```

