---
title: "Midterm 1"
author: "Matheus Ardigueri"
date: "2026-09-26"
output: html_document
---

# Exam Midterm CRSS 8010

# Question 1

```{r}
test.plot <- c("a","b")
_mydata <- c(1,2,3)
4varieties <- c(1,2,2)
.pseudoscore <- c(3,1,1)
Maize_42 <- c(1,1,1)
x <- c(1,1,1)
```

# Question 3

```{r}
j <- 1:20
k <- j

plot(j, k, type = "s")
```


# Question 4

```{r}
qf(0.01, df1 = 5, df2 = 30, lower.tail = FALSE)
```
# Question 7

```{r}
dof_treatments <- 3
dof_res <- 12
ssVar <- 21.47
ssRes <- 54.32

ms_trt <- ssVar / dof_treatments
ms_trt

ms_res <- ssRes / dof_res
ms_res

f_val <- ms_trt/ms_res
f_val

p_val <- pf( f_val, df1 = dof_treatments, df2 = dof_res, lower.tail = FALSE ) #df1 treatments, df2 residuals
p_val


```

```{r}
potato <- read.csv("../data/midterm1_potatoes.csv")

potato
```




# Question 18

```{r}
mean_grand_potatoes <- mean(potato$yield)

mean_sg <- mean(potato$yield[potato$variety == "Sugargate"])

ssq <- (mean_sg - mean_grand_potatoes) ^2 * 8 

ssq

```
# Question 19

```{r}
unique(potato$variety)
```
```{r}
model_potato <- aov(yield ~ variety + block, data = potato)

summary(model_potato)
```

```{r}
library(agricolae)

agricol <- HSD.test(model_potato, trt = "variety")
agricol
```








```{r}
```{r Another anova}

Fortune <- c(42, 47)
Bugz <- c(65, 62)
Eliminator <- c(48, 44)

scores <- c(Fortune, Bugz, Eliminator)

#FIrst step is to calcuulate the mean

meanF <- mean(Fortune)
meanB <- mean(Bugz)
meanE <- mean(Eliminator)
mean_all <- mean(scores)

# Second step is to calculate the sum of sqares

ssF <- (meanF - mean_all) ^2 * length(Fortune)
ssB <- (meanB - mean_all) ^2 * length(Bugz)
ssE <- (meanE - mean_all) ^2 * length(Eliminator)
ssall <- ssF + ssB + ssE

# Third step is to calculate the errors

errorF <- sum((Fortune - meanF)^2)
errorB <- sum((Bugz- meanB)^2)
errorE <- sum((Eliminator- meanE)^2)
error <- errorF + errorB + errorE

dof_treatments <- 2 
dof_error <- 3

ms_trt <- ssall / dof_treatments 
ms_error <- error / dof_error 

F <- ms_trt / ms_error

F


```




```{r}

# Oppening dataset

rcbd_mid <- read.csv("../data/Midterm1_rcbd.csv")
head(rcbd_mid, 5)

```
```{r}

# Sampling randomly inside Variety column

library(dplyr)

rcbd_new <- rcbd_mid %>%
  group_by(Block) %>%
  mutate(variety_new = sample(Variety))


rcbd_new
```

